A Constraint Solving Approach to Parikh Images of Regular Languages

### 2024-02-21 multiple.tex
typesetting
corrected flow
Fix lemma 6.1 re/side-conditions
Closes: #83

      % EXPAND
      \ExpandM & 
      \inferrule
      {
       {\begin{array}{c}
           \left\{
           \FlowEq(\Automaton_i, \Filter),~~
           \Connected(\Automaton_i, \Filter),~~
           \MonoidElement = \sum_{\Transition \in \Transitions_{\Automaton_i}} \Filter(\Transition) \cdot \Map(\Transition)
           \right\}_{i=1}^k,
           \\[4ex]
          \ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement}, ~
          \SomeInequalities, \SomeClause
        \end{array}}
        }
      {\ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause} & 
      None \\[5ex]
      \Materialise &
      \inferrule
      {{
         \begin{array}{c}
           \FlowEq(\Automaton', \Filter'),~
           \Connected(\Automaton', \Filter'),~
           \BindingSum(\Automaton_i, \Automaton_j, \Filter')
           \\\ImagePredicate{
                     \Tuple{\Automaton_1, \Automaton_{i-1}, \Automaton_{i+1}, \ldots,
                     \Automaton_{j-1}, \Automaton_{j+1}, \ldots, \Automaton_k,
                     \Automaton'}
           }{\Map}{\Filter'}{\MonoidElement},~ \SomeInequalities, \SomeClause
           \end{array}}
           }
      {\ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_i, \ldots, \Automaton_j, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement},~ \SomeInequalities, \SomeClause} &
      $\begin{array}{@{}l@{}}
         \Filter' = \Extend(\Automaton',\Filter),\\
         1 \leq i < j \leq k, \\
      \Automaton' =
         \Filter_{\SomeInequalities}(\Automaton_i) \times \Filter_{\SomeInequalities}(\Automaton_j)\\
      \end{array}$ \\
      \bottomrule
      \end{tabular}
      \caption{Additional derivation rules for products of arbitrarily many automata.}\label{tbl:rules:multi}
    \end{table}

@@

The multiplicity of transitions in the product then has to be related
to the multiplicities in the individual automata, modelled using the
$\BindingSum$ predicate. The predicate expresses that the multiplicity
of a transition~$\Transition \in \Transitions_{\Automaton_i}$ in
$\Automaton_i$ has to coincide with the sum of the multiplicities of
transitions in $\Filter_{\SomeInequalities}(\Automaton_i) \times \Filter_{\SomeInequalities}(\Automaton_j)$ derived from $t$,
and similarly for $\Automaton_j$:
%
  $$
  \BindingSum(\Automaton_i, \Automaton_j, \Filter') ~~=~~
  \begin{aligned}
  & \Set{ 
    \Filter'(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State, \State_R}}{\Label}{\Tuple{\State', \State'_R}}} \Filter'(\Transition')
  \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_i} } \cup\mbox{} \\ 
  & \Set{
    \Filter'(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State_L, \State}}{\Label}{\Tuple{\State'_L, \State'}}} \Filter'(\Transition') \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_j}
  }
  \end{aligned}
$$

@@

We extend \cref{eq:single-filter} from \cref{sec:single:example} with
definitions for $B$ shown in \cref{eq:multi-filter}:
\begin{equation}\label{eq:multi-filter}
    \begin{aligned}
        & \Filter(\FromLabelTo{A}{a / l_{B+}}{F_B} ) = \TransitionVar_6 
        & \Filter(\FromLabelTo{A}{c / l_{B+}}{A} ) = \TransitionVar_7 
        & \Filter(\FromLabelTo{B}{b / l_{B+}}{I_B}) = \TransitionVar_8 
        &\Filter(\FromLabelTo{B}{c / l_{B+}}{B}) = \TransitionVar_9  \\
        & \Filter(\FromLabelTo{B}{c / l_{B+}}{F_B})= \TransitionVar_{10}
        & \Filter(\FromLabelTo{I_B}{a / l_{B+}}{A})= \TransitionVar_{11}
        & \Filter(\FromLabelTo{I_B}{b / l_{B+}}{B})= \TransitionVar_{12}
        \FlowEq(\mathcal{B}, \Filter) &=  \begin{aligned}
            1 + \TransitionVar_{8}   &= \TransitionVar_{11} + \TransitionVar_{12}\\
            \land \TransitionVar_{11} &= \TransitionVar_6\\
            \land \TransitionVar_{12} &= \TransitionVar_8 + \TransitionVar_{10}
        \end{aligned}
    \end{aligned}
\end{equation}

@@

  This implies that also the rule~\ExpandM{} can only be applied
  finitely often since its side condition only allows applying it once
  per instance of an $\Image$ predicate containing a product, and only finitely
  many instances of $\Image$ can be introduced on each branch.
\end{proof}

>>>>

      % EXPAND
      \ExpandM & 
      \inferrule
      {
       {\begin{array}{c}
           \left\{
           \FlowEq(\Automaton_i, \Filter),~~
           \Connected(\Automaton_i, \Filter),~~
           \MonoidElement = \sum_{\Transition \in \Transitions_{\Automaton_i}} \Filter(\Transition) \cdot \Map(\Transition)
           \right\}_{i=1}^k,
           \\[4ex]
          \ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement}, ~
          \SomeInequalities, \SomeClause
        \end{array}}
        }
      {\ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement}, \SomeInequalities, \SomeClause} & 
      None \\[5ex]
      \Materialise &
      \inferrule
      {{
         \begin{array}{c}
           \FlowEq(\Automaton', \Filter'),~
           \Connected(\Automaton', \Filter'),~
           \BindingSum(\Automaton_i, \Automaton_j, \Automaton', \Filter')
           \\\ImagePredicate{
                     \Tuple{\Automaton_1, \Automaton_{i-1}, \Automaton_{i+1}, \ldots,
                     \Automaton_{j-1}, \Automaton_{j+1}, \ldots, \Automaton_k,
                     \Automaton'}
           }{\Map}{\Filter'}{\MonoidElement},~ \SomeInequalities, \SomeClause
           \end{array}}
           }
      {\ImagePredicate{\Tuple{\Automaton_1, \ldots, \Automaton_i, \ldots, \Automaton_j, \ldots, \Automaton_k}}{\Map}{\Filter}{\MonoidElement},~ \SomeInequalities, \SomeClause} &
      $\begin{array}{@{}l@{}}
         \Filter' = \Extend(\Automaton',\Filter),\\
         1 \leq i < j \leq k, \\
      \Automaton' =
         \Filter_{\SomeInequalities}(\Automaton_i) \times \Filter_{\SomeInequalities}(\Automaton_j)\\
      \end{array}$ \\
      \bottomrule
      \end{tabular}
      \caption{Additional derivation rules for products of arbitrarily many automata.}\label{tbl:rules:multi}
    \end{table}

@@

The multiplicity of transitions in the product then has to be related
to the multiplicities in the individual automata, modelled using the
$\BindingSum$ predicate. The predicate expresses that the multiplicity
of a transition~$\Transition \in \Transitions_{\Automaton_i}$ in
$\Automaton_i$ has to coincide with the sum of the multiplicities of
transitions in $\Filter_{\SomeInequalities}(\Automaton_i) \times \Filter_{\SomeInequalities}(\Automaton_j)$ derived from $t$,
and similarly for $\Automaton_j$:
%
  $$
  \BindingSum(\Automaton_i, \Automaton_j, \Automaton', \Filter') ~~=~~
  \begin{aligned}
  & \Set{ 
    \Filter'(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State, \State_R}}{\Label}{\Tuple{\State', \State'_R}} \in \Transitions_{\Automaton'}} \Filter'(\Transition')
  \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_i} } \cup\mbox{} \\ 
  & \Set{
    \Filter'(\Transition)  =  \sum\limits_{\Transition' = \FromLabelTo{\Tuple{\State_L, \State}}{\Label}{\Tuple{\State'_L, \State'}} \in \Transitions_{\Automaton'}} \Filter'(\Transition') \SuchThat \Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions_{\Automaton_j}
  }
  \end{aligned}
$$

@@

We extend \cref{eq:single-filter} from \cref{sec:single:example} with
definitions for $B$ shown in \cref{eq:multi-filter}:
\begin{equation}\label{eq:multi-filter}
  \begin{array}{rl@{\qquad}rl@{\qquad}rl}
     \Filter(\FromLabelTo{A}{a / l_{B+}}{F_B} ) &= \TransitionVar_6 
     &
     \Filter(\FromLabelTo{A}{c / l_{B+}}{A} ) &= \TransitionVar_7 
     &
      \Filter(\FromLabelTo{B}{b / l_{B+}}{I_B}) &= \TransitionVar_8 
     \\
     \Filter(\FromLabelTo{B}{c / l_{B+}}{B}) &= \TransitionVar_9 
     &
     \Filter(\FromLabelTo{B}{c / l_{B+}}{F_B}) &= \TransitionVar_{10}
     &
     \Filter(\FromLabelTo{I_B}{a / l_{B+}}{A}) &= \TransitionVar_{11}
     \\
      \Filter(\FromLabelTo{I_B}{b / l_{B+}}{B}) &= \TransitionVar_{12}
    \\[3ex]
% 
    \multicolumn{6}{c}{
        \FlowEq(\mathcal{B}, \Filter) =\quad  \begin{aligned}
            1 + \TransitionVar_{8}   &= \TransitionVar_{11} + \TransitionVar_{12} \wedge\mbox{}\\
            \TransitionVar_{11} &= \TransitionVar_6\wedge\mbox{}\\
            \TransitionVar_{12} &= \TransitionVar_8 + \TransitionVar_{10}\mbox{}\\
            \TransitionVar_6 + \TransitionVar_{10} &= 1
        \end{aligned}}
  \end{array}
\end{equation}

@@

  \ExpandM{} can only be applied precisely once per $\Image$~term since
  each application introduces an identical set of formulas and we have a generic
  side condition that no rule may add only redundant formulas.
\end{proof}
